### 2020-02-06 ncov.tex
prior discussion, some tiny tweaks

\documentclass[12pt]{article}
\usepackage[top=1in,left=1in, right = 1in, footskip=1in]{geometry}

\usepackage{graphicx}
%\usepackage{adjustbox}

@@

We reviewed 7 modeling studies of the novel coronavirus outbreak that were published online between January 23 and January 26, 2020 (Table 1).
Five studies \citep{liuncov, majumderncov, readncov, riouncov, zhaoncov} were uploaded to pre-print servers (bioRxiv, medRxiv, and SSRN); one report was posted on the website of Imperial College London \citep{imaincov}; and one report was posted on nextstrain.org \citep{bedfordncov}.
They vary widely in their statistical methods and in the amount of data used to infer \Ro.
\cite{bedfordncov}, \cite{imaincov}, and \cite{riouncov} simulated branching process models and compared the predicted number of cases from their models with the estimated number of total cases by January 18 (between 1,000 and 9,7000 \citep{imaincov0}).
\cite{readncov} fitted a deterministic, metapopulation Susceptible-Exposed-Infected-Recovered (SEIR) model to incidence data between January 1 and January 21 from major cities in China and other countries.
\cite{zhaoncov} fitted an exponential growth model to incidence data up to January 22 and inferred \Ro via the Euler-Lotka equation (\eref{euler}).
\cite{liuncov} modeled the observed incidence up to January 23 as a function of the observed incidence at previous time steps, the generation-interval distribution, and the basic reproductive number using a Poisson likelihood \citep{forsberg2008likelihood}.
Finally, \cite{majumderncov} fitted the Incidence Decay and Exponential Adjustment (IDEA) model \citep{fisman2013idea} to incidence data up to January 26, which is equivalent to fitting an exponential growth model and assuming a fixed generation-interval distribution.
Some of these studies have now been published in peer-reviewed journals \citep{riou2020pattern, zhao2020preliminary} or have been updated with better uncertainty quantification \citep{readncov2}.

@@

We use weakly informative priors on hyperparameters representing pooled means and noninformative priors on hyperparameters representing between-study standard deviations:
\begin{linenomath*}
\begin{equation}
\begin{aligned}
\mu_r &\sim \gammdist(\mathrm{mean}=1\,\mathrm{week}^{-1},\,\mathrm{shape}=2)\\
\mu_G &\sim \gammdist(\mathrm{mean}=1\,\mathrm{week},\,\mathrm{shape}=2)\\
\mu_\kappa &\sim \gammdist(\mathrm{mean}=0.5,\,\mathrm{shape}=2)\\
(\sigma_r, \sigma_G, \sigma_\kappa) &\sim \textrm{half-normal}(0, 10).
\end{aligned}
\end{equation}
\end{linenomath*}
Here, we followed recommendations outlined in \cite{gelman2006prior} by reparameterizing the hierarchical gamma model in terms of its mean and standard deviation and imposing a noninformative but proper prior distributions on between-study standard deviations, i.e., $\textrm{half-normal}(0, 10)$.
We had initially tried gamma priors with small shape parameters ($< 1$) on between-study shape parameters ($=\mu^2/\sigma^2$) of our hierarchical model but found this put too much on large between-study variances; this is in fact a known problem \citep{gelman2006prior}, and thus was a poor prior choice.
We also found that imposing strong priors on between-study shape parameters can be problematic as they \textit{a priori} assume large between-study variance, whereas imposing weak priors can allow for very low between-study variance and lead to poor mixing.

We run 4 independent Markov Chain Monte Carlo chains each consisting of 500,000 burnin steps and 500,000 sampling steps.
Posterior samples are thinned every 1000 steps.
Convergence is assessed by ensuring that the Gelman-Rubin statistic is below 1.01 for all hyperparameters \citep{gelman1992inference};
trace plots and marginal posterior distribution plots are presented in Appendix.
95\% confidence intervals are calculated by taking 2.5\% and 97.5\% quantiles from the marginal posterior distribution for each parameter.
\texttt{R} code is available in GitHub (https://github.com/parksw3/nCoV\_framework).

>>>>

\documentclass[12pt]{article}
\usepackage[top=1in,left=1in, right = 1in, footskip=1in]{geometry}
\usepackage{url}
\usepackage{graphicx}
%\usepackage{adjustbox}

@@

We reviewed 7 modeling studies of the novel coronavirus outbreak that were published online between January 23 and January 26, 2020 (Table 1).
Five studies \citep{liuncov, majumderncov, readncov, riouncov, zhaoncov} were uploaded to pre-print servers (bioRxiv, medRxiv, and SSRN); one report was posted on the web site of Imperial College London \citep{imaincov}; and one report was posted on nextstrain.org \citep{bedfordncov}.
They vary widely in their statistical methods and in the amount of data used to infer \Ro.
\cite{bedfordncov}, \cite{imaincov}, and \cite{riouncov} simulated branching process models and compared the predicted number of cases from their models with the estimated number of total cases by January 18 (between 1,000 and 9,7000 \citep{imaincov0}).
\cite{readncov} fitted a deterministic, metapopulation Susceptible-Exposed-Infected-Recovered (SEIR) model to incidence data between January 1 and January 21 from major cities in China and other countries.
\cite{zhaoncov} fitted an exponential growth model to incidence data up to January 22 and inferred \Ro via the Euler-Lotka equation (\eref{euler}).
\cite{liuncov} modeled the observed incidence up to January 23 as a function of the observed incidence at previous time steps, the generation-interval distribution, and the basic reproductive number using a Poisson likelihood \citep{forsberg2008likelihood}.
Finally, \cite{majumderncov} fitted the Incidence Decay and Exponential Adjustment (IDEA) model \citep{fisman2013idea} to incidence data up to January 26, which is equivalent to fitting an exponential growth model and assuming a fixed generation-interval distribution.
Some of these studies have now been published in peer-reviewed journals \citep{riou2020pattern, zhao2020preliminary} or have been updated with better uncertainty quantification \citep{readncov2}.

@@

We use weakly informative priors on hyperparameters representing pooled means and non-informative priors on hyperparameters representing between-study standard deviations:
\begin{linenomath*}
\begin{equation}
\begin{aligned}
\mu_r &\sim \gammdist(\mathrm{mean}=1\,\mathrm{week}^{-1},\,\mathrm{shape}=2)\\
\mu_G &\sim \gammdist(\mathrm{mean}=1\,\mathrm{week},\,\mathrm{shape}=2)\\
\mu_\kappa &\sim \gammdist(\mathrm{mean}=0.5,\,\mathrm{shape}=2)\\
(\sigma_r, \sigma_G, \sigma_\kappa) &\sim \textrm{half-normal}(0, 10).
\end{aligned}
\end{equation}
\end{linenomath*}
We followed recommendations outlined in \cite{gelman2006prior}, parameterizing the top-level gamma distributions in terms of their means and standard deviations and imposing weakly informative prior distributions on between-study standard deviations, i.e., $\textrm{half-normal}(0, 10)$.
We had initially used gamma priors with small shape parameters ($< 1$) on between-study shape parameters ($=\mu^2/\sigma^2$) but found this put too much prior probability on large between-study variances. This phenomenon is a known problem \citep{gelman2006prior}.
Alternative choices of prior for the between-study shape parameters are also suboptimal: imposing strong priors (e.g. half-Student-$t(\mu=0,\sigma=1,\nu=4)$  assumes \textit{a priori} that between-study variance is large,  while weak priors (e.g. half-Cauchy(0,5)) can lead to poor mixing.

We run 4 independent Markov Chain Monte Carlo chains each consisting of 500,000 burnin steps and 500,000 sampling steps.
Posterior samples are thinned every 1000 steps.
Convergence is assessed by ensuring that the Gelman-Rubin statistic is below 1.01 for all hyperparameters \citep{gelman1992inference};
trace plots and marginal posterior distribution plots are presented in Appendix.
95\% confidence intervals are calculated by taking 2.5\% and 97.5\% quantiles from the marginal posterior distribution for each parameter.
\texttt{R} code is available in GitHub (\url{https://github.com/parksw3/nCoV_framework}).
